Figures from: Vertebrate host availability and the thermal properties of mosquito-borne parasite transmission

### 2023-06-02 manuscript-figures.Rmd
add sensitivity figures

>>>>

```{r fig-SX_CT_distributions, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:CT_dists} The distributions of the critical thermal minimum, critical thermal maximum, and thermal optimum vary with vertebrate host availability.'}
all.density.df <- read_rds("results/CTminmaxTopt_densities.rds") %>%  
  ungroup() %>% 
  mutate(system_label = case_when(
    system_ID == "Aedes aegypti / DENV" ~ "(a) *Ae. aegypti* and DENV",
    system_ID == "Aedes aegypti / ZIKV" ~ "(b) *Ae. aegypti* and ZIKV",
    system_ID == "Aedes albopictus / DENV" ~ "(c&#41; *Ae. albopictus* and DENV",
    system_ID == "Anopheles gambiae / Plasmodium falciparum" ~ "(d) *An. gambiae* and *P. falciparum*",
    system_ID == "Culex quinquefasciatus / WNV" ~ "(e) *Cx. quinquefasciatus* and WNV"
  ))

### Option 1: inline violin plots ----
plot.densities_violin_inline <- ggplot() +
  geom_violin(data = filter(all.density.df, variable == "Thermal optimum"),
              aes(x = value, y = KH_factor, linetype = sigmaH, fill = "Topt")
  ) +
  geom_violin(data = filter(all.density.df, variable == "Critical thermal minimum"),
              aes(x = value, y = KH_factor, linetype = sigmaH, fill = "CTmin"),
  ) +
  geom_violin(data = filter(all.density.df, variable == "Critical thermal maximum"),
              aes(x = value, y = KH_factor, linetype = sigmaH, fill = "CTmax"),
  ) +
  scale_x_continuous(
    name = "Temperature (°C)",
    expand = c(0,0)
  ) +
  scale_y_discrete(
    name = "Vertebrate host population density (ind/ha)"
  ) +
  scale_linetype(
    name = "Biting tolerance\n(bites per host per day)"
  ) +
  scale_fill_discrete(
    name = "",
    breaks = c("CTmin", "Topt", "CTmax"),
    labels = c("Critical thermal minimum", "Thermal optimum", "Critical thermal maximum")
  ) +
  facet_wrap(~system_label, labeller = labeller()) +
  theme_minimal_grid(16) +
    theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    axis.title.x = element_text(hjust = 0.5),
    strip.text = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )
shift_legend(plot.densities_violin_inline)

### Option 2: inline box and whisker plots ----
plot.densities_boxwhisker_inline <- ggplot() +
  geom_boxplot(data = filter(all.density.df, variable == "Thermal optimum"),
               aes(x = value, y = KH_factor, linetype = sigmaH, fill = "Topt")
  ) +
  geom_boxplot(data = filter(all.density.df, variable == "Critical thermal minimum"),
               aes(x = value, y = KH_factor, linetype = sigmaH, fill = "CTmin"),
  ) +
  geom_boxplot(data = filter(all.density.df, variable == "Critical thermal maximum"),
               aes(x = value, y = KH_factor, linetype = sigmaH, fill = "CTmax"),
  ) +
  scale_x_continuous(
    name = "Temperature (°C)",
    expand = c(0,0)
  ) +
  scale_y_discrete(
    name = "Vertebrate host population density (ind/ha)"
  ) +
  scale_linetype(
    name = "Biting tolerance\n(bites per host per day)"
  ) +
  scale_fill_discrete(
    name = "",
    breaks = c("CTmin", "Topt", "CTmax"),
    labels = c("Critical thermal minimum", "Thermal optimum", "Critical thermal maximum")
  ) +
  facet_wrap(~system_label, labeller = labeller()) +
  theme_minimal_grid(16) +
    theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    axis.title.x = element_text(hjust = 0.5),
    strip.text = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )
shift_legend(plot.densities_boxwhisker_inline)

### Option 3: stacked violin plots ----
plot.densities_violin_stacked <- ggplot() +
  geom_violin(data = all.density.df,
              aes(x = value, y = KH_factor, fill = variable, linetype = sigmaH)) +
  scale_x_continuous(
    name = "Temperature (°C)",
    expand = c(0,0)
  ) +
  scale_y_discrete(
    name = "Vertebrate host population density (ind/ha)"
  ) +
  scale_linetype(
    name = "Biting tolerance\n(bites per host per day)"
  ) +
  scale_fill_discrete(
    name = ""
  ) +
  facet_wrap(~system_label, labeller = labeller()) +
  theme_minimal_grid(16) +
    theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    axis.title.x = element_text(hjust = 0.5),
    strip.text = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )
shift_legend(plot.densities_violin_stacked)

### Option 4: stacked box and whisker plots ----
plot.densities_boxwhisker_stacked <- ggplot() +
  geom_boxplot(data = all.density.df,
              aes(x = value, y = KH_factor, fill = variable, linetype = sigmaH)) +
  scale_x_continuous(
    name = "Temperature (°C)",
    expand = c(0,0)
  ) +
  scale_y_discrete(
    name = "Vertebrate host population density (ind/ha)"
  ) +
  scale_linetype(
    name = "Biting tolerance\n(bites per host per day)"
  ) +
  scale_fill_discrete(
    name = ""
  ) +  
  facet_wrap(~system_label, labeller = labeller()) +
  theme_minimal_grid(16) +
    theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    axis.title.x = element_text(hjust = 0.5),
    strip.text = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )
shift_legend(plot.densities_boxwhisker_stacked)

```{r fig-SX_R0_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:R0_sens} The sensitivity of the distribution of $\\mathcal{R}_0$ to mosquito life history traits.'}

R0.HPD <- read_rds("results/R0_HPD_sens.rds") %>%  
  ungroup() %>% 
  mutate(system_label = case_when(
    system_ID == "Aedes aegypti / DENV" ~ "(a) *Ae. aegypti* and DENV",
    system_ID == "Aedes aegypti / ZIKV" ~ "(b) *Ae. aegypti* and ZIKV",
    system_ID == "Aedes albopictus / DENV" ~ "(c&#41; *Ae. albopictus* and DENV",
    system_ID == "Anopheles gambiae / Plasmodium falciparum" ~ "(d) *An. gambiae* and *P. falciparum*",
    system_ID == "Culex quinquefasciatus / WNV" ~ "(e) *Cx. quinquefasciatus* and WNV"
  ))

# Label functions: 
appender_sigmaH <- function(string) {
  unname(TeX(paste("$\\sigma_H = $", string)))}
appender_KH <- function(string) {
  unname(TeX(paste("$K_H = $", string)))}

Temp_range_for_plot <- R0.HPD %>% 
  ungroup() %>%
  dplyr::filter(rel_HPD_width > eps) %>%
  select(Temperature) %>% range()

plot.R0.rel.sens <- R0.HPD %>% 
  filter(KH %in% c(1,100) & sigmaH %in% c(50, Inf)) %>% 
  filter(!(KH == 100 & sigmaH == Inf)) %>% 
  filter(rel_HPD_width != 0) %>% 
  # For each KH value, sum up all sensitivity values to get total sensitivity across parameters
  group_by(system_ID, sigmaH, KH, Temperature) %>% 
  mutate(sens_total = sum(rel_HPD_width)) %>% 
  # Divide each sensitivity value by the total sensitivity
  mutate(sens_rel = rel_HPD_width / sens_total) %>% 
  arrange(sens_rel) %>% 
  # Make a stacked plot, colored by parameter
  ggplot(aes(x = Temperature, y = sens_rel, fill = focal_var)) +
  geom_area() +
  scale_fill_discrete(
    name = "Focal parameter",
    breaks = c("betaV", "deltaL", "etaV", "muV", "rhoL", "sigmaV", "sigmaV_f"),
    labels =  c("Vector competence", "Pr(larval survival)", 
                "Parasite development rate", "Mortality rate", 
                "Immature development rate", "Max. biting rate", 
                "Eggs per female per day")
  ) +
  scale_x_continuous(
    limits = Temp_range_for_plot,
    expand = c(0,0)
  ) +
  scale_y_continuous(
    name = unname(TeX("Relative contribution to $R_0$ HPD width"))
  ) +
  facet_grid(cols = vars(system_label),
             rows = vars(KH, sigmaH), 
             labeller = labeller(
               system_label = label_wrap_gen(width = 10),
               sigmaH = as_labeller(appender_sigmaH, default = label_parsed),
               KH = as_labeller(appender_KH, default = label_parsed)),
             scales = "free_y") +
  theme_minimal_hgrid(16) +
    theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title.x = element_text(hjust = 0.5),
    strip.text.x = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )
plot.R0.rel.sens
# shift_legend(plot.R0.rel.sens)

```{r fig-SX_ddTR0_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:ddTR0_sens} The sensitivity of the distribution of the derivative of $\\mathcal{R}_0$ with respect to temperature to mosquito life history traits.'}

ddTR0.HPD <- read_rds("results/ddTR0_HPD_sens.rds") %>%  
  ungroup() %>% 
  mutate(system_label = case_when(
    system_ID == "Aedes aegypti / DENV" ~ "(a) *Ae. aegypti* and DENV",
    system_ID == "Aedes aegypti / ZIKV" ~ "(b) *Ae. aegypti* and ZIKV",
    system_ID == "Aedes albopictus / DENV" ~ "(c&#41; *Ae. albopictus* and DENV",
    system_ID == "Anopheles gambiae / Plasmodium falciparum" ~ "(d) *An. gambiae* and *P. falciparum*",
    system_ID == "Culex quinquefasciatus / WNV" ~ "(e) *Cx. quinquefasciatus* and WNV"
  ))

## Plot R0 uncertainty 
Temp_range_for_plot <- ddTR0.HPD %>% ungroup() %>% 
  dplyr::filter(HPD_width > eps) %>% 
  select(Temperature) %>%  range()

var_name_table <- list(
  betaV = c(TeX("$\\beta_V$")),
  deltaL = c(TeX("$\\delta_L$")),
  etaV = c(TeX("$\\eta_V$")),
  muV = c(TeX("$\\mu_V$")),
  rhoL = c(TeX("$\\rho_L$")),
  sigmaV = c(TeX("$\\sigma_V$")),
  sigmaV_f = c(TeX("$\\sigma_v f$"))
)

plot.ddTR0.rel.sens <- ddTR0.HPD %>% 
  filter(KH %in% c(1,100) & sigmaH %in% c(50, Inf)) %>% 
  filter(!(KH == 100 & sigmaH == Inf)) %>% 
  filter(rel_HPD_width != 0) %>% 
  # For each KH value, sum up all sensitivity values to get total sensitivity across parameters
  group_by(system_ID, sigmaH, KH, Temperature) %>% 
  mutate(sens_total = sum(rel_HPD_width)) %>% 
  # Divide each sensitivity value by the total sensitivity
  mutate(sens_rel = rel_HPD_width / sens_total) %>% 
  arrange(sens_rel) %>% 
  # Make a stacked plot, colored by parameter
  ggplot(aes(x = Temperature, y = sens_rel, fill = focal_var)) +
  geom_area() +
  scale_fill_discrete(
    name = "Focal parameter",
    breaks = c("betaV", "deltaL", "etaV", "muV", "rhoL", "sigmaV", "sigmaV_f"),
    labels =  c("Vector competence", "Pr(larval survival)", 
                "Parasite development rate", "Mortality rate", 
                "Immature development rate", "Max. biting rate", 
                "Eggs per female per day")
  ) +
  scale_x_continuous(
    limits = Temp_range_for_plot,
    expand = c(0,0)
  ) +
  scale_y_continuous(
    name = unname(TeX("Relative contribution to $\\frac{d}{dT}R_0$ HPD width"))
  ) +
  facet_grid(cols = vars(system_label),
             rows = vars(KH, sigmaH), 
             labeller = labeller(
               system_label = label_wrap_gen(width = 10),
               sigmaH = as_labeller(appender_sigmaH, default = label_parsed),
               KH = as_labeller(appender_KH, default = label_parsed)),
             scales = "free_y") +
  theme_minimal_hgrid(16) +
    theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title.x = element_text(hjust = 0.5),
    strip.text.x = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )
plot.ddTR0.rel.sens

```
```{r fig-SX_Topt_sensitivity, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:Topt_sens} The sensitivity of the thermal optimmum to mosquito life history traits as a function of vertebrate host population density when $\\sigma_H = 50$.'}
Topt.HPD <- read_rds("results/Topt_HPD_sens.rds") %>%  
  ungroup() %>% 
  mutate(system_label = case_when(
    system_ID == "Aedes aegypti / DENV" ~ "(a) *Ae. aegypti* and DENV",
    system_ID == "Aedes aegypti / ZIKV" ~ "(b) *Ae. aegypti* and ZIKV",
    system_ID == "Aedes albopictus / DENV" ~ "(c&#41; *Ae. albopictus* and DENV",
    system_ID == "Anopheles gambiae / Plasmodium falciparum" ~ "(d) *An. gambiae* and *P. falciparum*",
    system_ID == "Culex quinquefasciatus / WNV" ~ "(e) *Cx. quinquefasciatus* and WNV"
  )) %>%
  filter(!is.na(rel_HPD_width)) %>%
  # For each KH value, sum up all sensitivity values to get total sensitivity
  group_by(system_ID, sigmaH, KH) %>% 
  mutate(sens_total = sum(rel_HPD_width)) %>% 
  # Divide each sensitivity value by the total sensitivity
  mutate(sens_rel = rel_HPD_width / sens_total) #%>% 
  # arrange(sens_rel)

# !!! mention the value of sigmaH

plot.Topt.rel.sens <- Topt.HPD %>% 
  # Make a stacked plot, colored by parameter
  ggplot(aes(x = KH, y = sens_rel, fill = focal_var)) +
  geom_area() +
  scale_fill_discrete(
    name = "Focal parameter",
    breaks = c("betaV", "deltaL", "etaV", "muV", "rhoL", "sigmaV", "sigmaV_f"),
    labels =  c("Vector competence", "Pr(larval survival)", 
                "Parasite development rate", "Mortality rate", 
                "Immature development rate", "Max. biting rate", 
                "Eggs per female per day")
  )  +
  scale_x_log10(
    name = TeX("Vertebrate host population density (ind/ha)"),
    expand = c(0, 0),
    limits = c(0.05, 1E3),
    breaks = 10^seq(-2, 4, 1),
    labels = TeX(c("$0.01$", "$0.1$", "$1$", "$10", "$100$", "$10^3$", "$10^4$"))
  ) +
  scale_y_continuous(
    name = unname(TeX("Relative contribution to $T_{opt}$ HPD width"))
  ) +
  # scale_y_continuous(
  #   limits = c(0,1)
  # ) +
  facet_wrap(~system_label, scales = "free") +
  theme_minimal_hgrid(16) +
    theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title.x = element_text(hjust = 0.5),
    strip.text.x = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )
shift_legend(plot.Topt.rel.sens)
